Objective-first optimization for nanophotonics

### 2012-05-31 presentations/meta_book/my_chapter.tex
Getting started on text.

% Some custom commands.
\newcommand{\BE}{\begin{equation}}
\newcommand{\EE}{\end{equation}}
\newcommand{\BA}{\begin{eqnarray}}
\newcommand{\EA}{\end{eqnarray}}
\newcommand{\curl}{\nabla\times}
\newcommand{\minimize}[1]{\JLUminimize_{#1}\;&}
\newcommand{\subto}{\text{subject to}\;&}
\newcommand{\myfig}[2]{\begin{figure}[!h]\includegraphics[width=\textwidth]{fig/#1}\caption{#2}\label{#1}\end{figure}}

\chapter{Objective-First Nanophotonic Design}
\label{intro}

@@

Introductory paragraph.

\input{wave_eq}
\input{ob1}
\input{1d}
\input{res}
\input{wg}

>>>>

% Some custom commands.
\newcommand{\BE}{\begin{equation}}
\newcommand{\EE}{\end{equation}}
\newcommand{\BA}{\begin{eqnarray}}
\newcommand{\EA}{\end{eqnarray}}
\newcommand{\curl}{\nabla\times}
\newcommand{\minimize}[1]{\JLUminimize_{#1}\;&}
\newcommand{\subto}{\text{subject to}\;&}
\newcommand{\myfig}[2]{\begin{figure}[!h]\includegraphics[width=\textwidth]{fig/#1}\caption{#2}\label{#1}\end{figure}}
\newcommand{\BI}{\begin{itemize}\item}
\renewcommand{\I}{\item}
\newcommand{\EI}{\end{itemize}}
\newcommand{\ER}[1]{\eqref{eq:#1}}
\chapter{Objective-First Nanophotonic Design}
\label{intro}

@@

In this chapter, we introduce an ``objective-first'' strategy 
    for designing nanophotonic devices.

General description:
\BI Only uses desired fields.
\I  Can attempt to design any linear device.
\I  Allows for non-physical fields. \EI

This strategy is unique in that it
    asks the user only for what electromagnetic field
    the device should produce,
    and then attempts to generate the design
    without further user intervention.

\input{wave_eq}
\input{ob1}
% \input{1d}
% \input{res}
% \input{wg}

### 2012-05-31 presentations/meta_book/ob1.tex
Getting started on text.

\section{The objective-first design problem}
intro

\subsection{Design objectives}
Talk about $f(x)$ and that we are interested in convex ones.

>>>>

\section{The objective-first design problem}
We now build off of the field-solver and the structure-solver,
    as previously outlined,
    by formulating the design problem
    and outlining the objective-first strategy.

\subsection{Design objectives}
A design objective, $f(x)$, is simply defined as
    a function we wish to be minimal 
    for the design to be produced.

For instance, in the design of a device
    which must transmit efficiently into a particular mode,
    we could choose $f(x)$ to be the negative power into that mode.
Or, if the device was to be a low-loss resonator,
    we could choose $f(x)$ to be the amount of power leaking 
    out of the device.

### 2012-05-31 presentations/meta_book/wave_eq.tex
Getting started on text.

\section{The electromagnetic wave equation}
\subsection{Physics formulation}
Let's talk about the electromagnetic wave equation from a physics standpoint.
Let's start from Maxwell's equations without currents.
\BA \curl E = - \mu_0 \frac{\partial H}{\partial t} \\
    \curl H = J + \epsilon \frac{\partial E}{\partial t} \EA

\BA \curl E = - i \mu_0 \omega H \\
    \curl H = J + i \epsilon \omega E \EA

\BE \curl \epsilon^{-1} \curl H - \mu_0 \omega^2 H = \curl \epsilon^{-1} J \EE

Talk about reduced dimensions here. Refer reader to appendix.

\subsection{Numerical formulation}
Now let's talk about the electromagnetic wave equation from a computational perspective.

To solve the wave equation on a computer we need to use the Yee grid.

To make things easier we will use weird units.

We also need to take care of boundary conditions.

\subsection{Solving for $H$}
The wave equation from a mathematical perspective.

\BE (\curl \epsilon^{-1} \curl - \mu_0 \omega^2) H = \curl \epsilon^{-1} J \EE

Becomes, with a change of variables.

\BE A(p)x = b(p) \EE

This can be solved directly in 1D and 2D. 
Special methods needed in 3D. 
We just do 1D and 2D.

\subsection{Solving for $\epsilon^{-1}$}
The wave equation from a optimization perspective.

Because scalar multiplication is transitive 
    ($\epsilon^{-1} (\curl H) = (\curl H) \epsilon^{-1}$ and
    $\epsilon^{-1} J = J \epsilon^{-1}$)
\BE \curl (\curl H) \epsilon^{-1} - \curl J \epsilon^{-1}  = \mu_0 \omega^2 H  \EE

which we write as

\BE B(x)p = d(x) \EE

Special constraints... (binary)

\subsection{Insight}
Basically, we see that the electromagnetic wave equation is separably linear in $H$ and $\epsilon^{-1}$

This means that...

>>>>

\section{The electromagnetic wave equation}

We start by understanding the underlying physical equation.
We extend it beyond its usual use in simulation, 
    and use it for design.

\subsection{Physics formulation}
First, let's derive our wave equation,
    starting with the differential form of Maxwell's equations, 
\BA \curl E = - \mu_0 \frac{\partial H}{\partial t} \\
    \curl H = J + \epsilon \frac{\partial E}{\partial t}, \EA
    where $E$, $H$, and $J$ are 
    the electric, magnetic and electric current % TODO: Check this.
    vector fields, respectively,
    $\epsilon$ is the permittivity
    and $\mu_0$ is the permeability, which we assume to be 
    that of vacuum everywhere.

For time dependence $\exp(-i \omega t)$, 
    where $\omega$ is the angular frequency,
    these become
\BA \curl E = - i \mu_0 \omega H \\
    \curl H = J + i \epsilon \omega E, \label{eq:H2E} \EA
    which we can combine to form our wave equation,
\BE \curl \epsilon^{-1} \curl H - \mu_0 \omega^2 H = \curl \epsilon^{-1} J. 
    \label{eq:wave} \EE
For further information,
    as well as simplifications to the wave equation in reduced dimensions,
    please see the Appendix. % TODO: Use reference.

\subsection{Numerical formulation}
Now, on top of the analytical formulation of the wave equation \ER{wave}
    we will now add a numerical, or discretized, formulation.
This will be needed in order to solve for arbitrary structures.

First, we discretize our computational space 
    according to the Yee grid, % TODO: Reference!
    which allows us to easily define the curl $(\curl)$
    operators in \ER{wave} as described in the Appendix. % TODO: ref.
This allows us, with a change of variables to formulate \ER{wave} as
\BE A(p)x = b(p), \label{eq:Ab} \EE 
    where $H \to x$, $\epsilon^{-1} \to p$; 
    and where
\BE A(p) = \curl \epsilon^{-1} \curl - \mu_0 \omega^2 \EE 
    and
\BE b(p) = \curl \epsilon^{-1} J. \EE
Note that our use of $A(p)$ and $b(p)$ instead of $A$ and $b$
    simply serves to clarify the dependence
    of both $A$ and $b$ to $p$.

Additionally,  we use periodic boundary conditions
    with stretched-coordinate % TODO: ref.
    perfectly matched layers where necessary for our examples.

\subsection{Solving for $H$}
With our numerical formulation, we can now solve for the $H$-field
    (the $E$-field can be computed from the $H$-field using \ER{H2E})
    by using general linear algebra solvers.
Doing so is also simply known as a time-harmonic or 
    a finite-difference frequency-domain (FDFD) simulation.

Now, while a full three-dimensional problem
    is computationally quite taxing;
    in one- and two-dimensions, 
    \ER{Ab} is easily solved using the standard sparse solver
    included in Matlab,
    and this technique is regularly employed 
    in the examples which follow.

\subsection{Solving for $\epsilon^{-1}$}
The next step,
    after having built a field-solver or simulator
    (finds $x$ given $p$) for our wave equation,
    is to build a structure-solver for it.
In other words, we need to be able to solve for $p$ given $x$.

To do so, we return to \ER{wave}
    and remark that 
    $\epsilon^{-1} (\curl H) = (\curl H) \epsilon^{-1}$ and
    $\epsilon^{-1} J = J \epsilon^{-1}$ 
    since scalar multiplication is communicative.
This allows us to rearrange \ER{wave} as
\BE \curl (\curl H) \epsilon^{-1} - \curl J \epsilon^{-1}  = \mu_0 \omega^2 H  \EE
which we now write as 
\BE B(x)p = d(x), \label{eq:Bd} \EE 
    where
\BE B(x) = \curl (\curl H) - \curl J\EE
    and 
\BE d(x)  = \mu_0 \omega^2 H.  \EE

Solving this system would now seem to allow us
    to choose an electromagnetic field
    and then find the structure to produce it;
    which strongly suggests that it will be useful in
    the design of nanophotonic devices.

In terms of computational complexity, as with \ER{Ab}, 
    \ER{Bd} in its current form can be solved using standard tools.

\subsection{Bi-linearity of the wave equation}
Although additional mathematical machinery must still be added
    in order to get a useful design tool,
    we have really shown so far is that the wave equation is 
    separately linear or \emph{bi-linear} in $x$ and $p$.
Namely that,
\BE A(p)x-b(p) = B(x)p - d(x). \EE
In other words, fixing $p$ makes solving the wave equation for $x$
    a linear problem, and vice versa.
Note that the joint problem,
    where both $x$ and $p$ are allowed to vary,
    is not linear.

The bi-linearity of the wave equation
    is fundamental in our objective-first strategy
    which relies on the fact
    that we already know how to solve linear systems well,
    and is the reason why we chose $\epsilon^{-1} \to p$
    instead of the more natual $\epsilon \to p$.
Indeed, this property forms a natural division of labor
    in the objective-first scheme, which we outline below.
